---
title: "clabsi_data_analysis"
author: "Jiayi Yang"
date: "2022-10-04"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(dplyr)
library(ggridges)
```

Importing and tidying CLABSI data, adding a period of 14 days

```{r}
test = 
  central_lined_df =
  read_excel("./CLABSI_data.xlsx", sheet = 1, range = cell_cols("A:G")) %>% 
  arrange(EMPI) %>% 
  select(EMPI, ordering_date) %>% 
mutate(
  status = "central_lined",
  end_date = ymd(ordering_date) + days(14)
) %>% 
  distinct(EMPI, ordering_date, .keep_all = TRUE) %>% 
  mutate(
      period_1 = interval(ymd(ordering_date), ymd(end_date)),
      date = ordering_date
  ) %>% 
  select(EMPI, status, date, period_1) %>% 
  arrange(EMPI,date) 

test2 = 
  test[!duplicated(test$EMPI), ]
test2
```
We have `r nrow(central_lined_df)` patients who have central lines inserted in our sample

Bloodstream Infectious importing and cleaning
```{r}
bcp_df =
  read_excel("./CLABSI_data.xlsx", sheet = 3, range = cell_cols("A:K")) %>% 
  arrange(EMPI) %>% 
  select(EMPI, PROC_NAME, RESULT_TIME) %>% 
  filter(PROC_NAME == "BLOOD CULTURE") %>% 
  mutate(
  status = "blood_culture_positive",
  date = as.Date(RESULT_TIME, "%y/%m/%d")
)  %>% 
  distinct(EMPI, date, .keep_all = TRUE)
bcp_df
```

We have `r nrow(bcp_df)` patients who have blood culture positive shown in the EMR in our sample

Next, see if result_time of BCP aligns with central line insert period.
Trying to join central_lined_df and bcp_df to see patients who are cl inserted and bcp
```{r}
join_test=
join_cl_bcp =
  left_join(test2, bcp_df, by = c("EMPI")) %>%
  mutate(
    bcp_status = ifelse(date.y %within% period_1, 1, 0)
  ) %>%    
drop_na() %>% 
  select(EMPI, bcp_status, date.x, date.y) %>% 
  mutate(
    duration = date.y - as.Date(date.x)) %>% 
  arrange(EMPI, desc(bcp_status))
  
join_test2 = join_test[!duplicated(join_test$EMPI), ]

join_test2
```
Table it BCP within 14 days
```{r}
  knitr::kable(join_test2)
```


Great, we found `r nrow(join_cl_bcp)` observations of patients who have blood culture positive within the 14-day period of central line inserted. 

Incidence of bcp in central lined patients are `r nrow(join_cl_bcp)` / `r nrow(central_lined_df)`.


make an EDA table
```{r}
EDA_plot =
  left_join(central_lined_df, join_cl_bcp, by = c("EMPI")) %>% 
  drop_na()
EDA_plot
```


Next, we look at TPN data
```{r}
tpn_df =
   read_excel("./CLABSI_data.xlsx", sheet = 4, range = cell_cols("A:K")) %>% 
  arrange(EMPI) %>% 
  select(EMPI, START_DATE, END_DATE) %>% 
  mutate(
    status = "tpn"
  )
tpn_df
```
```{r}

distr_plot = 
  tpn_df %>% 
    group_by(EMPI) %>% 
 summarize(n_obs = n()) %>% 
   mutate(patient = 1:n(),
          patient = as.integer(patient))
ggplot(distr_plot, aes(x = n_obs)) +
  geom_histogram() + scale_x_continuous(breaks = seq(0, 150, by = 10)) + scale_y_continuous(breaks = seq(0, 40, by = 2)) +
  labs(
    title = "Distribution of PN days for patients",
    x = "# of days on PN",
    y = "# of patients were on PN for that many days"
  ) 
distr_plot


```
##```{r}
ggplot(distr_plot, aes(x = patient, y = n_obs)) + 
  geom_point(aes(color = "patient"), alpha = .5) +
  geom_line(na.rm = FALSE) +
  labs(
    title = "Distribution of PN days for each patient",
    x = "patient",
    y = "PN days observations"
  ) + 
    theme(legend.position = "bottom")
##```

## R Markdown
