---
title: "New Data Analysis"
author: "Jiayi Yang"
date: "2022-11-16"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(dplyr)
library(ggridges)
library(ggplot2)
```
```{r}
library("MatchIt")
library(survival)
```


import the new data
```{r, message=FALSE}
new_data = 
  read_excel("./LDA.xlsx", sheet = 1, range = cell_cols("A:K")) %>% 
  filter(flo_meas_name %in% c(  'G IP NYC LDA MIDLINE DOUBLE LUMEN',
                                'G IP NYC LDA MIDLINE SINGLE LUMEN',
                                'G LDA PULMONARY ARTERY CATHETER/SWAN GANZ',
                                'LDA CVC QUADRUPLE LUMEN',
                                'LDA CVC SINGLE LUMEN',
                                'LDA CVC TRIPLE LUMEN',
                                'LDA DUAL LUMEN PORT',
                                'LDA EXTERNAL URINARY CATHETER',
                                'LDA PICC SINGLE LUMEN',
                                'LDA PICC TRIPLE LUMEN',
                                'LDA PUMP DEVICE',
                                'NYC LDA PRESSURE INJURY',
                                'RETIRED - NYC ADT TC PICC DOUBLE LUMEN')
         ) %>% 
  mutate(
    placement_date = as.Date(PLACEMENT_INSTANT, "%y/%m/%d"),
    removal_date = as.Date(REMOVAL_INSTANT, "%y/%m/%d"),
    duration = removal_date - placement_date
  ) %>% 
  filter(removal_date != '2157-11-19')

new_data
new_data2 = new_data[!duplicated(new_data$EMPI), ] 
```

We have `r nrow(new_data)` patients who have central lines inserted in our sample

```{r}
bcp_df =
  read_excel("./CLABSI_data.xlsx", sheet = 3, range = cell_cols("A:K")) %>% 
  arrange(EMPI) %>% 
  select(EMPI, PROC_NAME, RESULT_TIME) %>% 
  filter(PROC_NAME == "BLOOD CULTURE") %>% 
  mutate(
  status = "blood_culture_positive",
  date = as.Date(RESULT_TIME, "%y/%m/%d")
)  %>% 
  distinct(EMPI, date, .keep_all = TRUE)
bcp_df
```
We have `r nrow(bcp_df)` patients who have blood culture positive shown in the EMR in our sample

Next, see if result_time of BCP aligns with central line insert period.
Trying to join central_lined_df and bcp_df to see patients who are cl inserted and bcp during their central line placement period.

```{r}
join_test=
join_cl_bcp =
  left_join(new_data, bcp_df, by = c("EMPI")) %>%
  mutate(
   period_1 = interval(ymd(placement_date), ymd(removal_date)),
   cl_duration_time = difftime(ymd(removal_date), ymd(placement_date), units="days"),
    bcp_status = ifelse(date %within% period_1, 1, 0)
  ) %>%    

  select(EMPI, bcp_status, date, placement_date, removal_date, period_1, cl_duration_time) %>% 
  arrange(EMPI, desc(bcp_status))
  
join_test2 = join_test[!duplicated(join_test$EMPI), ] %>% 
  mutate(
    matching_period = date - placement_date,
    new_bcp_status = ifelse(is.na(bcp_status), 0, bcp_status)
  )

table(join_test2$new_bcp_status)
```

Great, we found 595 observations of patients who have blood culture positive within their periods of central line inserted. 

The incidence for blood culture positive in central lined patients are 595/6754 = 8.8%

Now we try to match controls to cases by first full join the bcp_status table (join_test2) with our original new_data table:
```{r}
full_table = 
  full_join(new_data, join_test2, by = c("EMPI")) 
full_table2 = full_table[!duplicated(full_table$EMPI), ] 

full_table2$bcp_status[is.na(full_table2$new_bcp_status)] <- 0

table(full_table2$new_bcp_status)
```
We have 595 cases and 1840 unique controls to pool from. `full_table2` is the full table until now.

Matching by variable `matching_period` which stands for the time between central line placement and tested blood culture positive

by MatchIt:
```{r}
full_table3 = full_table2 %>% 
  mutate(
    new_matching_period = ifelse(is.na(matching_period), 0, matching_period)
  )

set.seed(1234)
match.it <- matchit(new_bcp_status ~ new_matching_period, data = full_table3, method="nearest", ratio=1)
matched_df <- summary(match.it)

knitr::kable(matched_df$nn, digits = 2, align = 'c', 
      caption = 'Table 2: Sample sizes')

```
We have 595 cases and 595 matched controls by their matching period(time of central line inserted to time of bcp). There are 1245 remained unmatched, the total of the sample population is 2435.

save the matched dataset into a new dataframe named `df.match`
```{r}
df.match <- match.data(match.it)[1:ncol(full_table3)]
rm(matched_df)

df.match %>% 
  mutate(
    BIRTH_DATE = as.Date(BIRTH_DATE, "%y/%m/%d"),
    sex_c = case_when(sex_c == 1 ~ "Male",
              sex_c == 2 ~ "Female") %>% 
      as.factor()
  ) 
df.match
```
df.match has 1190 observations include 595 cases and 595 controls


demographic analysis: age, gender
1 is female, 2 is male.
...

```{r}

ggplot(aes(x = sex_c), data=subset(df.match, !is.na(sex_c)), color = sex_c) +
  geom_histogram(binwidth=30) + labs(
    title = "Distribution of gender",
    x = "gender",
    y = "count"
  ) +
  facet_grid(. ~sex_c) 

```

Look at the exposure TPN status
Next, we look at TPN data
```{r}
tpn_df =
   read_excel("./CLABSI_data.xlsx", sheet = 4, range = cell_cols("A:K")) %>% 
  arrange(EMPI) %>% 
  select(EMPI, START_DATE, END_DATE) %>% 
  mutate(
    status = "tpn"
  )
join_tpn = 
  left_join(df.match, tpn_df, by = c("EMPI")) %>% 
  mutate(
    tpn_status = ifelse(START_DATE %within% period_1, 1, 0)
  ) %>% 
    arrange(EMPI, desc(tpn_status)) %>% 
  select(EMPI, BIRTH_DATE, sex_c, new_bcp_status, period_1, START_DATE, END_DATE, tpn_status, duration, hosp_admsn_time, new_matching_period) %>% 
  mutate(
    new_tpn_status = ifelse(is.na(tpn_status), 0, tpn_status)
  ) %>% 
  mutate(
    birth_date = as.Date(gsub(BIRTH_DATE,pattern="00:00:00",replacement="",fixed=T)),
    hosp_admsn_date = as.Date(gsub(hosp_admsn_time,pattern="00:00:00",replacement="",fixed=T)),
    age = year(hosp_admsn_date) - year(birth_date)
  ) %>% 
    distinct(EMPI, .keep_all = TRUE) 

tpn_table = join_tpn%>% 
    janitor::tabyl(new_tpn_status, new_bcp_status)
tpn_table
```
This is the 2*2 table of exposure on the left, outcome on the right.
OR = ad/bc = 0.8969

## Conditional logistic regression

Possible covariates: age (age at the hospital admission time point), gender (sex_c coded as 1 for female, 2 for male, 3 is unknown), central line placement duration (duration)...

First, using `mylogit` to do a logistic regression on the outcome variable (having blood culture positive) and on the predictor variable (TPN status). We also want to look at potential confounder so we included duration as a covariate.
```{r}
fit.clr <- clogit(new_bcp_status ~ new_tpn_status + duration + age + sex_c + strata(new_matching_period), data = join_tpn)

round(summary(fit.clr)$coef, 4)
```

```{r}
OR.CI <- cbind("OR" = exp(coef(fit.clr)),
               exp(confint(fit.clr)))
round(OR.CI, 3)
```

Test result discussion:
For the exposure predictor `new_tpn_status`, p-value 0.2743 is greater than 0.05 which indicates that it is not statistically significant. Our model suggests that having total parental nutrition does not significantly impact the result of blood culture positive. The Odds Ratio of having TPN to BCP is eβ = e^-0.5393 = 1.715.
This indicates that: the odds of having blood culture positive in the total parental nutrition group is 1.715 times the odds of having blood culture positive in the non total parental nutrition group

For the predictor `duration`, p-value 0.0001 is smaller than 0.05 which indicates that it is statistically significant. Our model suggests that duration of central lines placement does significantly impact the result of blood culture positive. The Odds Ratio of central line placement duration to BCP is eβ = e^0.0153 = 1.015.
This indicates that: An increase of 1 day in central line placement period is associated with an increase of 1.5% in the odds of blood culture positive, adjusting for tpn, age and gender.

?For the predictor `age`, p-value 0.0559 is slightly bigger than 0.05 which indicates that it is not statistically significant. Our model suggests that the age of patient does not significantly impact the result of blood culture positive. The exponentiated coefficient for age variable in patients of getting BCP is eβ = e^-0.0085 = 0.992.
This indicates that: For every one year increase in age, the odds of central-lined patients getting blood culture positive decrease by 0.8%, adjusting for tpn, duration and gender.

For the predictor `sex_c`, p-value 0.9284 is greater than 0.05 which indicates that it is not statistically significant. Our model suggests that gender does significantly impact the result of blood culture positive. The Odds Ratio of being a female and getting BCP compares to being a male and getting BCP is eβ = e^0.0166 = 1.017.
Gender is not a significant covariate in the relationship of getting blood culture positive. 


### Regular Logistic Regression Model (will not use)
```{r}
mylogit <- glm(new_bcp_status ~  duration + age + sex_c, data = join_tpn, family = "binomial")
```
```{r}
summary(mylogit)
```
```{r}
confint.default(mylogit)
```



------------

Verification of assumptions, model diagnostics, forward selection model
Discussion, final evaluation

May look at the interaction term for age and gender?
normality assumption
report the r square : how much the variability 

